---
title: " "
format:
  html: 
    df-print: kable
    code-fold: show
    code-summary: "Hide code"
    code-overflow: wrap
    toc-title: Page Contents
    toc: true
    toc-depth: 2
    toc-location: right
    number-sections: false
    html-math-method: katex
    smooth-scroll: true
editor: source
editor_options: 
  chunk_output_type: console
---
```{=html}
<style type="text/css">

body, td {
   font-size: 13pt;
}
code.r{
  font-size: 9pt;
}
pre {
  font-size: 11pt
}
</style>
```

::: {.callout-note icon=true}
#### VSV Sequences    
Sequence data has been submitted to [GenBank](https://www.ncbi.nlm.nih.gov/genbank/) but may not yet been released due to a pending data report publication still in review.  Code below will provide an accession numbers table.  This page will be updated when genomic data becomes publicly available.  
:::


#### Download *sequence accession numbers* 
{{< downloadthis assets/genome_sequences.csv dname=genome_sequences label="Get Table" icon=download type=info class=data-button >}}    **Sequence accession numbers** (*22 kb, csv*) 

```{r message=FALSE, warning=FALSE, echo=FALSE}
#Load needed R-packages to execute code

library(tidyverse)
library(here)
```

### Read Data    
Read sequence accession numbers from archived file on [Open Science Framework](https://osf.io/ghzfq/).  Also available as supplemental data with publication.    
```{r message=FALSE, warning=FALSE}

seq_accessions <- read_csv("https://osf.io/download/zytwj/")

```

```{r}
seq_accessions
```